RadLib: a radiative property model library for combustion CFD

### 2021-07-19 submissions/CPC_response_to_reviewers.md

----------------------------------------------------

## Reviewer 2

A radiation spectral property library is developed in this study using C++. A Planck-mean absorption coefficient based model, a WSGG model and a RCSLM model from the literature are implemented  in the package. Verification results are presented and discussed. Although the authors addressed an important issue of radiation modeling, the contribution is marginal in the reviewer's opinion. No new models/methods are proposed in this study. As the authors correctly pointed out, the solver for radiative transfer equation is equally important, which is not included in the library. The coupling of a spectral model to a specific RTE solver, can have different levels of challenges depending on the combination. The design of the library doesn't really address any of the RTE-related issues, nor provide any new solutions to simplify the coupling of spectral models to RTEs. If the authors can demonstrate that the library can be coupled with popular CFD software and popular RTE solvers such as P1 and DOM, it would further improve the quality of this study. A few other concerns are listed as follows:

**The authors would like to thank the reviewer for their comments and suggestions. It is true that the library presented does not develop new models, but implementing and providing those models is nontrivial, and we believe makes a good contribution that is within the scope of this software journal.**

**We agree the reviewer's suggestion regarding coupling with popular CFD software and a corresponding RTE solver, and that this would strengthen the paper. We have made significant revisions to the paper and code in this regard. First, we developed a new Fortran interface that calls the underlying C++ code. Second, we coupled the code to the Fire Dynamics Simulator (FDS) code, developed by NIST. FDS is a Fortran code that required the new Fortran interface. FDS has several radiation models available including WSGG DOM implementation. Radlib was interfaced to this model by using the radlib fortran module in the source code. We then replaced the existing calls to the functions for computing the absorption coefficient and weighting factors with calls to radlib. FDS does an outer loop over the gases/bands with inner loops over the local grid points. RadLib was not set up for this, since it computes the properties for all bands at a given point. Hence, direct use of RadLib would have been inefficient. As a result, we implemented a new interface function get_k_a_oneband, that facilitated implementation in FDS. In the revised paper, we present results comparing the WSGG and RCSLW models in FDS applied to the FM Burner validation case that is detailed in the FDS Validation Documentation. We hope that this effort will satisfy the reviewer and improve the paper as desired.**

@@

1. RADCAL data were used to generate the PlanckMean absorption coefficient. Based on the reviewer's experience, the PlanckMean absorption coefficients can have meaningful differences when generated from RADCAL and from HiTEMP2010. It is recommended a more recent database being used for the library.

	**The Planck Mean absorption coefficients model as implemented in RadLib refers specifically to the radiation model in [3,5], which is one of the most common radiation models applied to turbulent combustion simulations. It was chosen because of its popularity and widespread use in combustion simulations (despite its limitations). That said, there is value in using the most current tools. Given the reviewer's experience with variance in the Planck Mean absorption coefficients as calculated from various databases, the authors will strongly consider adding an updated Planck Mean model using the most current available data. That said, the PM model implemented is generally applied in optically-thin situations with relatively low radiative transfer in which case the errors are not as significant.**

@@

2. Page 10, missing information on [INSERT RUNTIME COMPARISONS HERE].

	**This was a placeholder statement that should not have been included in the document. It has been removed.**

3. In Figure 2, the RCSLM seems to consistently perform better than WSGG, except in Example S2. In addition, WSGG performs worse than PM in Example S4. Can the authors explain why?

	**The PM and WSGG models are primarily empirical and contain very little representative physics, which could explain their unexpected behavior in Examples 2 and 4, but also makes it very difficult to find or treat sources of error. Any model, but especially empirical models, can behave unexpectedly when applied to conditions or parameters outside of those for which it was developed. We were suprised by the poor performance of the WSGG model in the case the author notes. This is one area where it is nice to have a model (as presented) that is easily implemented with access to validation cases that can be extended. The example cases are meant to be illusrative of the software's capabilities, not an evaluation of the models themselves, which is outside the scope of this paper. The reviewer's request for an explanation, however, is valid. The discussion in Section 5 has been expanded and revised to address these specific discrepancies. Additional information about the specific limitations, strengths, and weaknesses associated with each model can be found in Section 2.**

@@

4. Equation 6 appears to be inaccurate.

	**Upon review, the authors find no errors in Equation 6, which matches Equation 20.65 of the general band model (of which the WSGG model is a subset) derivation in [2] and appears to be consistent with the other equations in the paper. The equation is q = int(I(s)sdΩ). The notation here is that I is a function of direction, but is itself a scalar so that I(s)s is equivalent to a vector I in direction s.**

@@

5. Typo: page 4, line 49, "relating them" -> "relating the"

	**This phrase has been removed in the revison.**

-------------------------------------

### 2021-07-20 manuscript/manuscript.tex
DOL: adding FDS results for the FM burner

@@

Figure~\ref{f:cost}a displays the relative cost to compute the gas properties $k$ and $a$, represented by the time required to execute the C++ function \texttt{get\textunderscore k\textunderscore a()}, averaged over one million evaluations. The results for each model are normalized by the WSGG model, which required an average of 8.0E-7~s. Relative to the WSGG model, the cost to evaluate the gas properties with the PM model was 0.098. The RCSLW model was evaluated with $n=1$, 2, 4, and 8 gray gases, which resulted in relative costs of 1.9, 4.9, 9.4, and 18, respectively. The cost of the RCSLW model is very nearly proportional to the number of gray gases considered.
%
\begin{figure}
    \begin{center}
    \begin{tabular}{c c}
        (a) & (b) \\
        \includegraphics[width=2.75 in]{fig_getka_c++.pdf} &
        \includegraphics[width=2.75 in]{fig_exS3_c++.pdf}
    \end{tabular}
    \caption{Relative computational cost to (a) evaluate gas properties only and (b) run Example S3, including the ray-tracing solver. Results are normalized by the cost of the WSSG model for each case.}
    \label{f:cost}
    \end{center}
\end{figure}
%

>>>>

\subsection{CFD Results} \label{s:cfd}

The RadLib model was coupled to the Fire Dynamics Simulator (FDS) code developed at NIST \cite{FDS}. FDS is written in Fortran, and the Fortran interface of Radlib was used for simulations. FDS includes several radiative models, but the spectral WSGG model coupled to a finite volume discrete ordinates model is used here. This was done by editing the \texttt{radi.f90} file. A \texttt{use rad\textunderscore module} statement is included and the functions \texttt{A\textunderscore WSGG} and \texttt{KAPPA\textunderscore WSGG} were replaced with calls to RadLib's \texttt{get\textunderscore k\textunderscore a\textunderscore oneband}. The \texttt{oneband} version was used because FDS does an outer iteration over the bands/gases, computing properties at a given band for each grid point.

FDS includes a number of validation cases. Here, we compare to the FM Burner case. The configuration is a vertically fired gas burner with various fuels and with a coflow oxidizer stream of varying nitrogen dilution. The burner has outer and inner diameters of 15.2 and 13.7 cm and is near the floor in the center of a 1.22 by 1.22 by 1.83 m tall compartment. The heat release rate was 10 kW and included a 1 kW pilot in a surrounding ring. The coflow oxidizer was fed through the floor. Simulations performed here used the ethylene fuel with 20.9\% \ce{N2} oxidizer (i.e., air). Further FDS model and case simulation details are provided in the FDS validation and user guides \cite{FDS}.

Figure~\ref{f:fds} shows results of the simulation with comparison to experimental data. Three radiation models were included: the FDS default model, and the RCSLW and WSGG models from RadLib. 
The FDS default model uses a single gray gas with a composition and temperature dependent absorption coefficient computed from RADCAL. Further details are provided in the FDS user guide \cite{FDS}. 
Figure~\ref{f:fds}a shows radial profiles of mean temperature at two vertical positions. The models produce centerline temperature differences of about 80 K (7\%) between the RCSLW and the WSGG at z=2.5D, and 90 K (12\%) at z=3.5D. The differences are lower at lower points in the flame since the effect of radiation on temperature is cumulative. Figure~\ref{f:fds}b plots height versus radiative emission per unit length along the centerline with the spread in radiative emission around the peak values being 1.3 kW/m (11\%).
%
\begin{figure}
    \begin{center}
    \begin{tabular}{c c}
        (a) & (b) \\
        \includegraphics[width=2.5 in]{fig_fds_T_mean.pdf} &
        \includegraphics[width=2.5 in]{fig_fds_rad_mean.pdf}
    \end{tabular}
    \caption{Radial profiles of mean temperature at two heights (a), and height versus radiative emission (b) for FDS simulations of the FM Burner validation case comparing three models to experimental data.}
    \label{f:fds}
    \end{center}
\end{figure}
%

@@

Figure~\ref{f:cost}a displays the relative cost to compute the gas properties $k$ and $a$, represented by the time required to execute the C++ function \texttt{get\textunderscore k\textunderscore a()}, averaged over one million evaluations. The results for each model are normalized by the WSGG model, which required an average of 8.0E-7~s. Relative to the WSGG model, the cost to evaluate the gas properties with the PM model was 0.098. The RCSLW model was evaluated with $n=1$, 2, 4, and 8 gray gases, which resulted in relative costs of 1.9, 4.9, 9.4, and 18, respectively. The cost of the RCSLW model is very nearly proportional to the number of gray gases considered.
%
\begin{figure}
    \begin{center}
    \begin{tabular}{c c}
        (a) & (b) \\
        \includegraphics[width=2.5 in]{fig_getka_c++.pdf} &
        \includegraphics[width=2.5 in]{fig_exS3_c++.pdf}
    \end{tabular}
    \caption{Relative computational cost to (a) evaluate gas properties only and (b) run Example S3, including the ray-tracing solver. Results are normalized by the cost of the WSSG model for each case.}
    \label{f:cost}
    \end{center}
\end{figure}
%
